preciseTAD: A machine learning framework for precise TAD boundary prediction at base-level resolution

### 2020-08-26 Manuscript/manuscript.Rmd
cleanup ???

We assessed the biological significance of our predicted boundaries by their association with several known molecular drivers of 3D chromatin architecture, including CTCF, RAD21, SMC3, and ZNF143 ChIP-Seq peaks. To do so, we used the *computeMatrix*, *plotProfile*, and *plotHeatmap* tools offered by deepTools (version 2.0) [@ramirez2016deeptools] <!--via integration with Galaxy (version 3.3.2.0.0) [@afgan2018galaxy]--> [??? Correct, change to new pipeline; !!!removed galaxy usage]. Additionally, we compared the median $log_2$ genomic distances between TAD boundaries and the same top predictive ChIP-seq annotations using Wilcoxon Rank-Sum tests. Furthermore, we compared the overlap between predicted and called boundaries in GM12878 and K562 cell lines. Boundaries were first flanked by resolution, *r*, and overlaps were visualized using Venn diagrams from the `Vennerable` R package (version 3.1.0). Overlaps were further quantified using the Jaccard index defined as

@@

*preciseTAD* offers flexibility in controlling both the number of predicted boundaries and the distance between them. The two primary parameters are the probability threshold $t$ and $\epsilon$ (referred to as eps; parameter of DBSCAN). The combination of these two quantities changes the resulting number of predicted boundaries from *preciseTAD*. Lower values of $t$ and $\epsilon$ will result in more clusters of bases, and therefore, more boundaries. As a heuristic, we evaluated the pairwise combination of four different thresholds ($t=(0.95,0.975,0.99,1.0)$) and four different $eps$ values ($\epsilon=(r, 2r, 3r, 4r, 5r)$)), where $r=5000$, the optimal resolution. We found that the normalized overlaps - calculated as the total number of ChIP-seq peaks that overlapped within a given flanked boundary, divided by the number of boundaries - between top TFBS sites and flanked *preciseTAD* boundaries were maximized for combinations of $t=1.0$ and $eps=r$ (Supplementary Figure 10). [??? We talked about relaxing t threshold, but it requires extensive reruns. Where are we on that?]

>>>>

We assessed the biological significance of our predicted boundaries by their association with several known molecular drivers of 3D chromatin architecture, including CTCF, RAD21, SMC3, and ZNF143 ChIP-Seq peaks. To do so, we used the *computeMatrix*, *plotProfile*, and *plotHeatmap* tools offered by deepTools (version 2.0) [@ramirez2016deeptools]. Additionally, we compared the median $log_2$ genomic distances between TAD boundaries and the same top predictive ChIP-seq annotations using Wilcoxon Rank-Sum tests. Furthermore, we compared the overlap between predicted and called boundaries in GM12878 and K562 cell lines. Boundaries were first flanked by resolution, *r*, and overlaps were visualized using Venn diagrams from the `Vennerable` R package (version 3.1.0). Overlaps were further quantified using the Jaccard index defined as

@@

*preciseTAD* offers flexibility in controlling both the number of predicted boundaries and the distance between them. The two primary parameters are the probability threshold $t$ and $\epsilon$ (referred to as eps; parameter of DBSCAN). The combination of these two quantities changes the resulting number of predicted boundaries from *preciseTAD*. Lower values of $t$ and $\epsilon$ will result in more clusters of bases, and therefore, more boundaries. As a heuristic, we evaluated the pairwise combination of four different thresholds ($t=(0.95,0.975,0.99,1.0)$) and four different $eps$ values ($\epsilon=(r, 2r, 3r, 4r, 5r)$)), where $r=5000$, the optimal resolution. We found that the normalized overlaps - calculated as the total number of ChIP-seq peaks that overlapped within a given flanked boundary, divided by the number of boundaries - between top TFBS sites and flanked *preciseTAD* boundaries were maximized for combinations of $t=1.0$ and $eps=r$ (Supplementary Figure 10).
